---
output: github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/"
)
```

# EpiSoon

[![Travis-CI Build Status](https://travis-ci.org/epiforecasts/EpiSoon.svg?branch=master)](https://travis-ci.org/epiforecasts/EpiSoon)

*Warning: This package is a work in progress and is currently developed solely with the COVID-19 outbreak in mind. Breaking changes may occur and the authors cannot guarantee support.*

**Aim:** To forecast the time-varying reproduction number and using this to forecast reported case counts. 

## Installation

Install the analysis and all dependencies with: 

```{r, eval = FALSE}
remotes::install_github("epiforecasts/EpiSoon", dependencies = TRUE)
```


## Quick start


* Load packages (`bsts` for models, `ggplot2` for plotting, and `cowplot` for theming)

```{r, message = FALSE}
library(EpiSoon)
library(bsts)
library(future)
library(cowplot)
library(dplyr)
```

* Set up example data (using `EpiSoon::example_obs_rts` and `EpiSoon::example_obs_cases` as starting data sets). When generating timeseries with `EpiNow` use `get_timeseries` to extract the required data.

```{r}
obs_rts <- EpiSoon::example_obs_rts %>%
   dplyr::mutate(timeseries = "Region 1") %>%
   dplyr::bind_rows(EpiSoon::example_obs_rts %>%
  dplyr::mutate(timeseries = "Region 2"))

obs_cases <- EpiSoon::example_obs_cases %>%
   dplyr::mutate(timeseries = "Region 1") %>%
   dplyr::bind_rows(EpiSoon::example_obs_cases %>%
   dplyr::mutate(timeseries = "Region 2"))
```

* Define the list of `bsts` models to be compared.

```{r}
models <- list("AR 3" =
                  function(ss, y){bsts::AddAr(ss, y = y, lags = 3)},
               "Semi-local linear trend" =
                  function(ss, y){bsts::AddSemilocalLinearTrend(ss, y = y)})
```

* Compare models across timeseries (change the `future::plan` to do this in parallel).

```{r}
future::plan("sequential")

## Compare models
forecasts <- compare_timeseries(obs_rts, obs_cases, models,
                                   horizon = 7, samples = 10,
                                   serial_interval = EpiSoon::example_serial_interval)

forecasts
```

* Plot an evaluation of Rt forecasts using iterative fitting.

```{r, fig.width = 7, fig.height = 7, dpi = 320, out.width = "60%"}
plot_forecast_evaluation(forecasts$forecast_rts, obs_rts, c(7)) +
   ggplot2::facet_grid(model ~ timeseries) +
   cowplot::panel_border()
```

* Plot an evaluation of case forecasts using iterative fitting


```{r, fig.width = 7, fig.height = 7, dpi = 320, out.width = "60%"}
plot_forecast_evaluation(forecasts$forecast_cases, obs_cases, c(7)) +
   ggplot2::facet_grid(model ~ timeseries, scales = "free") +
   cowplot::panel_border()
```

* Summarise the forecasts by model scored against observed cases

```{r}
summarise_scores(forecasts$case_scores)
```


## Docker

This package was developed in a docker container based on the `rocker/geospatial` docker image. 

To build the docker image run (from the `EpiSoon` directory):

```{bash, eval = FALSE}
docker build . -t episoon
```

To run the docker image run:

```{bash, eval = FALSE}
docker run -d -p 8787:8787 --name episoon -e USER=episoon -e PASSWORD=episoon episoon
```

The rstudio client can be found on port :8787 at your local machines ip. The default username:password is epinow:epinow, set the user with -e USER=username, and the password with - e PASSWORD=newpasswordhere. The default is to save the analysis files into the user directory.

To mount a folder (from your current working directory - here assumed to be `tmp`) in the docker container to your local system use the following in the above docker run command (as given mounts the whole `episoon` directory to `tmp`).

```{bash, eval = FALSE}
--mount type=bind,source=$(pwd)/tmp,target=/home/EpiSoon
```

To access the command line run the following:

```{bash, eval = FALSE}
docker exec -ti episoon bash
```
