---
title: "Getting started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  dpi = 320,
  fig.height = 6,
  fig.width = 8
)
```


## Introduction

This vignette briefly outlines the functionality of `EpiSoon`. To get started load the required packages.


* Load the package (`bsts` for models, `ggplot2` for plotting, and `cowplot` for theming)

```{r, message = FALSE}
library(EpiSoon)
library(bsts)
library(cowplot)
library(dplyr)
```


## Forecast Rts, score and plot


* We use an example dataframe built into the package but this could be replaced with your own data.

```{r}
EpiSoon::example_obs_rts
```

* Fit a `bsts` model and produce a Rt forecast. Us appropriately wrapped any model can be used (see `bsts_model` for an example wrapper).
```{r}
rt_forecast <- forecast_rt(EpiSoon::example_obs_rts[1:10, ],
                       model = function(...){EpiSoon::bsts_model(model = function(ss, y){bsts::AddAutoAr(ss, y = y, lags = 10)}, ...)},
                       horizon = 21, samples = 10)

rt_forecast
```

* Score the forecast

```{r}
rt_scores <- score_forecast(rt_forecast, EpiSoon::example_obs_rts)

rt_scores
```

* Summarise the forecast scores

```{r}
summarise_scores(rt_scores)
```

* Summarise the forecast

```{r}
summarised_rt_forecast <- summarise_forecast(rt_forecast)

summarised_rt_forecast
```

* Plot the forecast against observed data

```{r}
plot_forecast(summarised_rt_forecast, EpiSoon::example_obs_rts)
```

## Forecast cases, score and plot

* Forecasting cases requires the observed cases on which the observed Rt estimates were based

```{r}
EpiSoon::example_obs_cases
```

* It also requires an assumption to be made about the serial interval (defined using probability distribution).

```{r}
EpiSoon::example_serial_interval
```

* Forecast cases (using the case data on which the observed Rt estimates were based)

```{r}
case_forecast <- forecast_cases(EpiSoon::example_obs_cases, rt_forecast,
                                serial_interval = EpiSoon::example_serial_interval)

case_forecast
```


* Score the cases forecast

```{r}
case_scores <- score_case_forecast(case_forecast, EpiSoon::example_obs_cases)

case_scores
```

* Summarise the cases scores

```{r}
summarise_scores(case_scores)
```

* Summarise the cases forecast

```{r}
summarised_case_forecast <- summarise_case_forecast(case_forecast)

summarised_case_forecast
```

* Plot the forecast against observed case data

```{r}
plot_forecast(summarised_case_forecast, EpiSoon::example_obs_cases)
```

## Use iterative fitting to explore a forecast

* To explore the quality of a models forecast it can help to iteratively forecast from each available data point. This is supported in `EpiSoon` using the following:

```{r}
it_rt_forecast <- iterative_rt_forecast(EpiSoon::example_obs_rts,
                                        model = function(...){
                                          EpiSoon::bsts_model(model = function(ss, y){bsts::AddAutoAr(ss, y = y, lags = 10)}, ...)
                                          },
                                        horizon = 7, samples = 10, min_points = 4)

it_rt_forecast
```


* We can then iteratively forecast cases using the following:

```{r}
it_cases_forecast <- iterative_case_forecast(it_fit_samples = it_rt_forecast,
                                             cases = EpiSoon::example_obs_cases,
                                             serial_interval = EpiSoon::example_serial_interval)

it_cases_forecast
```

* All functionality shown above is also supported for iterative forecasting.

## Evaluate a model

In real world use we are likely to want to evaluate a model by iteratively forecasting Rts and cases, summarising these forecasts, scoring them and then returning them in a sensible format. These steps are all contained in the `evaluate_model` function.

```{r}
model_eval <- evaluate_model(EpiSoon::example_obs_rts,
                             EpiSoon::example_obs_cases,
                             model = function(...){
                                          EpiSoon::bsts_model(model = function(ss, y){bsts::AddAutoAr(ss, y = y, lags = 10)}, ...)
                                          },
                             horizon = 21, samples = 10,
                             serial_interval = EpiSoon::example_serial_interval)

model_eval
```
* All functionality outlined above can be applied to this output but a special plotting function (`plot_forecast_evaluation`) is also provided. First evaluate the Rt forecast against observed values.

```{r}
plot_forecast_evaluation(model_eval$forecast_rts,
                         EpiSoon::example_obs_rts,
                         horizon_to_plot = 7)
```

* Then evaluate forecast cases against observed values.


```{r}
plot_forecast_evaluation(model_eval$forecast_cases,
                         EpiSoon::example_obs_cases,
                         horizon_to_plot = 7)
```

## Wrapper functions

`EpiSoon` provides several wrapper functions (`compare_models` and `compare_timeseries`). These both wrap `evaluate_model` and can be used to rapidly explore several forecasting models (`compare_models`) against multiple time series (`compare_timeseries`). All lower level summary and plotting functions can be then used with the output of these wrappers to explore the results. See the function documentation for further details.

